% !TEX root = ../notes.tex

\section{First Order PDEs}

We are going to consider a parabolic PDE,
$$ \pd u t = \pdd u x $$
with an initial condition $u(x, 0) = u_0(x)$. Consider a uniform set of time steps $t_m = m\Delta t$ for $m = 0, 1,2 \dots$. We also consider $x_j = a + j\Delta x$, where $\Delta x = (b-a)/N$. We write $u_j^m = u(x_j, t_m)$ and we seek to approximate $u_j^m$ by $U_j^m$ for $j = 0, 1, 2 \dots, N$ and $m = 0, 1, 2, \dots$. We write,
$$ \pdd u x (x_j, t) = \frac{u(x_{j+1}, t) - 2u(x_j, t) + u(x_{j-1}, t)}{\Delta x^2} + \mathcal{O}(\Delta x^2). $$
Similarly, we can do fortward and backward differences,
$$ \pd u t (x, t_m) = \frac{u(x, t_{m+1}) - u(x, t_{m})}{\Delta t} + \mathcal{O}(\Delta t), $$
and similar for a backward difference. Then forward,
$$ \frac{U^{m+1}_j - U_j^m}{\Delta t} = \frac{U_{j+1}^m - 2U_j^m + U_{j-1}^m}{\Delta x^2}, $$
and similarly for the others. Then we can discretise the boundary conditions.\\

\noindent
For ODEs, forward Euler was easy to implement. This kind of holds here. However, for implicit we have a system of non-linear equations. Let's consider forward euler first though,
$$ U_j^{m+1} = U_j^m + \mu (U_{j+1}^m - 2U_j^m + U_{j-1}^m) $$
Thus once we have the initial and boundary data we can find our starting points. Then we have all the $U_j^1$ and then all the $U_j^2$ and so on. We now let,
$$ \d_x^2 U_j = U_{j+1} - 2U_j + U_{j-1}. $$
Then the $\theta$-method becomes,
$$ \frac{U_{j}^{m+1} - U_j^m}{\Delta t} = \theta \frac{\d_x^2 U_j^{m+1}}{\Delta x^2} + (1 - \theta) \frac{\d^2 _x U_j^m}{\Delta x^2} $$
and it becomes a massive mess after some rearranging. We are going to set $m = 0$. We can write,
$$ (I - \mu \theta A)U^{m+1} = (I' + \mu (1 - \theta)A)U^m + g^{m+1} $$
where $U^m$ is the vector of the $U_j^m$'s, then $I'$ is the identity minus the top and bottom $1$'s are zero, and $g^{m+1} = (u_a(t_{m+1}), 0, \dots, 0, u_n(t_{m+1}))^T$.

\noindent
If we have a BC of $\a u + \b \pd u t = \g$. Then we can just take a forward difference given $x_0 = a$,
$$ \frac{u(x_1, t) - u(x_0, t)}{\Delta x} + \mathcal{O}(\Delta x) $$
This means,
$$ \a U_0^{m+1} + \b \frac{U_1^{m+1} - U_0^{m+1}}{\Delta x} = \g $$
and then in Explicit Euler, we have something more complicated. It is harder to approximate the solution on the boundary. We do something similar with the $\theta$-method. Just rearrange and fiddle our equations to no longer have upper line of all zeros. However, this is not $\Delta x$, so we want to get a second order finite difference scheme for our BC. Hence,
$$ \pd u t (a, t) = \frac{u(x_1, t) - u(x_{-1}, t)}{2\Delta x} + \mathcal{O}(\Delta x^2)$$
Hence we have,
$$ \a U_0^{m+1} + \b \frac{U_1^{m+1} - U_{-1}^{m+1}}{2\Delta x} = \g, $$
and we rearrange for $U_{-1}^{m+1}$ and put in when appropriate.