% !TEX root = ../notes.tex

What if $A_{11} = 0$? For example $A = \begin{pmatrix}
  0 & 1 \\ 1 & 0
\end{pmatrix}$.  We should be able to do this! Solution, pivot. This reorders the rows of $A_i$. So instead of,
$$ A_i = \ell_i u_i^T + \begin{pmatrix}
  & 0 & \\
  0 &*&* \\
  &*&*
\end{pmatrix} $$
we aim for,
$$ P_1A = \ell_i u_i^T + \begin{pmatrix}
  & 0 & \\
  0 &*&* \\
  &*&*
\end{pmatrix} $$
where $P_i$ is a permutation matrix.  For example, $P_i = \begin{pmatrix}
  0 &&& 1 \\
  &&1& 0 \\
  & \ddots && \\ %ugh rotate `\ddots`
  1 &&&
\end{pmatrix}$
this says that $P_iA$ exchanges the rows and $AP_i$ exchanges columns. We are going to use these $P_i$'s to just move the largest non-zero entries to the top. So instead of what we expect to have, we have,
$$ \tilde P_1 P_0 A = \begin{pmatrix}
  1 & 0 \\
  0 & P_1
\end{pmatrix} \left(\ell_1 u_1^T + \begin{pmatrix}
   & 0 \\
   0 & A_1
\end{pmatrix}\right) $$
where, $\tilde P_1 =  \begin{pmatrix}
  1 & 0 \\
  0 & P_1
\end{pmatrix}$, and so we redefine, $\ell_1 = \tilde P_{n-1} \dots \tilde P_0 \ell_1$ and so we now have,
$$ \tilde P_{n-1} \dots \tilde P_0 = \tilde \ell_1 u_1^T + \dots \tilde \ell_n u_n^T  $$
but does this hurt our structure? Unfortunately not. This is because each $\tilde P_n$ only acts on the structure of the part we want it to. Hence we have we $LU$ factoriation,
$$ A = \begin{pmatrix}
  0&&& \\
  \ell_1 & 0 && \\
  \vdots&\ell_2 &0& \\
  &\vdots& \ddots & \\
  &&& \ell_n
\end{pmatrix} $$

\subsection{Cholesky Factorisation}
We have $A = A^T \succ 0$, we have $\mathrm{eigs}(A) > 0$. We are going to force $L = U$ throughout the operation. Hence,
$$ A = \ell_1\ell_1^T + \begin{pmatrix}
  & 0 \\
  0 & A_1
\end{pmatrix} $$
We can say that $A_{11} > 0$ and that, $A_1 \succ 0$. Then RECURSION! Hence we have, after $n$ steps, we have the Cholesky, $R^TR$. We recall the $A^TA$ Gram Matrix. If we have some indefinite matrix, when $A = A^*$ we have $A = LDL^*$ where $L$ is triangular, but $D$ is a block matrix.

\subsection{QR Factoriation}
Consider some over-determined problem, $Ax = b$ usually not possible. Then we want to $\min_x \norm{Ax - b}_2$. To solve, $A = LU$ is useless. We need something different, we need the $QR$ factorisation. We let $A = QR$, where $Q^TQ = I$ and $R$ is upper triangular. This exists for any $A$. \\

\noindent
